A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
fix patterns

This section presents an overview of the usage of \eval in our corpus.
Richards et al. reported 82\% of JavaScript-enabled web pages using
\eval~\cite{ecoop11}. The case for R is more striking with 100\% of the
programs call the \eval function.\footnote{We combine calls to \c{eval},
  \c{eval.parent}, \c{evalq} and \c{local}.}  This high number is explained
by the fact that the R implementation, its core packages, use \eval
extensively. We differentiate uses based on their origin. A {\bf core} \eval
is one that occurs in the code of a core package.  A {\bf dependent} \eval
occurs in a package imported by the target package. An {\bf own} \eval is
one that occurs in the source code of the target package.

@@

%% SORRY  I HAD TO REFORMAT
\begin{table}[tb] \centering
 \begin{tabular}{r|c|c|c} \hline
Source & Core    & Dependent                      & Own \\\hline
CRAN   & 5.6M &  867K    & 259K \\
       & (1 / 337 / 46K)& (0 / 52 / 58K ) & (0 / 15 / 15K) \\\hline
Kaggle &  1.2M &  470K   & 0\\
 &  (0 / 721 / 18K) & (0 / 290 / 20K) & (0 / 0 / 0)\\\hline
\end{tabular}\caption{Eval calls in CRAN and Kaggle}\label{A}\end{table}

Table~\ref{table:site-package-summary} shows the number of call sites in
CRAN packages. \OneCallSitesEnrich packages have a single call site, and one package, \emph{metafor}, has \MaxCallSitesEnrich
call sites of \eval. \NoCallSitesEnrich packages have no calls to \eval. The core packages
have \NbCoreEvalCallSites call sites of \eval and Kaggle has none. These numbers are
under-approximations as \eval may be called reflectively, through aliases,
or passed to a higher-order function.

@@

In this section we turn our attention to the expression passed for evaluation to
\eval. The \eval function accepts all types of R objects as inputs, but in the
majority of cases (\AllExpressionInputEvalCallPerc), it is called with
expressions. The remaining \AllValueInputEvalCallPerc are values, \ie, they
evaluate to themselves.
%
%Furthermore, the proportion of values passed to core
%\eval call is \missingTag{CoreEvalValueInputPerc} compared to a significantly higher
%\missingTag{PackageEvalValueInputPerc} for package \eval calls.
%
Passing a value into an eval is effectively an no-op so we look closely at the
expression types the call sites pass to eval across all the runs.
\AllValueInputEvalSitePerc of all call sites always pass values,
\AllExpressionInputEvalSitePerc pass expressions and a very large proportion,
\AllPolymorphicInputEvalSitePerc, pass both expressions and values as inputs
across all program runs. We inspected the eval callsites that accept only values
or both values and expression as inputs. The top three of these call sites
contribute to over ~80\% of all calls to evals. The first one is inside
\c{match.arg} from core. Given a vector of names as strings, this function finds
the default values for parameters with those names in the caller function. Since
the default value for a function parameter can be a complex expression or a
benign value such as \c{NULL}, \c{match.arg} evaluates it through a call to
\eval. This makes the \eval inside \c{match.arg} ``polymorhpic''. The second
such \eval call originates from the popular \ggplot package of R which is a DSL
for plotting. \ggplot has it's own object system, \ggproto, which uses \eval
inside the \ggproto function to determine the super-class from which it
inherits. The default value of the superclass is \c{NULL} but in general could
be an arbitrarily complex expression. The third \eval call is in core R \c{str}
function. This function is used to print a summary of R objects. It internally
maps \eval function on the bindings of an environment which results in inputs of
multiples types being passed to \eval.

@@

We looked at the top ten expressions passed to core and package \eval calls. The
most frequent ten expressions to eval calls from core R contribute to 85.2\% of
all \eval calls.
\begin{table*}[tb]
  \centering
  \begin{tabular}{l|rr} \hline
    Expression & Eval Call & Eval Call \% \\\hline
    \c{c("auto", "shell", "radix")} & 1987105 & 29.3\%\\
    \c{c("auto", "shell", "quick", "radix")} & 1593169  & 23.5\%\\
    \c{\{info <- loadingNamespaceInfo(...} & 1008632 &       14.9\%\\
    \c{c("onLoad", "attach", "detach", "onUnload")}   & 470566 &      6.9\%\\
    \c{c("append", "prepend", "replace")} &              261587&       3.9\% \\
    \c{c("left", "right", "centre", "none")} & 162086     & 2.4\%\\
    \c{c("no", "ifany", "always")}   &                71580 &       1.1\%\\
    \c{c("pearson", "kendall", "spearman")}  & 72962 &      1.1\%\\
    \c{NULL}& 75330  &      1.1\% \\
    \c{Symbol}&                 66279&       1\%\\\hline
  \end{tabular}
  \caption{Top ten eval calls in Core}
  \label{A}
\end{table*}

@@

The most frequent ten expressions to eval calls from CRAN packages contribute to
77.1\% of all \eval calls.
\begin{table*}[tb]
  \centering
  \begin{tabular}{l|rr} \hline
    Expression & Eval Call & Eval Call \% \\\hline
    \c{Environment} &                                                              989302   & 61.8\%\\
    \c{column[rows] <<- what} &                                                    55677    & 3.5\%\\
    \c{function(value) freduce(value, `_function_list`)} &                         37251    & 2.3\%\\
    \c{NULL} &                         32005    & 2\%\\
    \c{List} &                         22293    & 1.4\%\\
    \c{c("default", "default2012", "default2011" ...}&                             20610    & 1.3\%\\
    \c{force(..1)}                                        &                        20461    & 1.3\%\\
    \c{alist(`_spec`)}                                   &                         18532    & 1.2\%\\
    \c{inner}                                           &                          18530    & 1.2\%\\
    \c{String Vector}                           &                           17487     & 1.1\%\\
  \end{tabular}
  \caption{Top ten eval calls in CRAN}
  \label{A}
\end{table*}

The expression \c{Environment} occurs because of the four callsites explained
above, \c{ggplot2::ggproto}, \c{R6::generator_funs}, \c{future::backtrace} and
\c{RModel::str.RMmodel}. The next expression, \c{column[rows] <<- what}, is used
inside the \c{plyr::rbind.fill} function to merge data frames by assigning
concatenated vectors to rows. The \c{<<-} operator is interesting in that it
skips the current scope and assigns in a parent scope in which the variable is
already present. In our corpus, all these \eval calls contribute to a single
side-effect. The expression \c{function(value) freduce(value, `_function_list`)}
arises from the \c{magrittr::\%>\%} function which is a pipe operator that pipes
the output of previous command to the next one. The expression is evaluated in a
custom environment to create a function binding for evaluating the components of
the pipe.\c{String Vector} and \c{List} also arise from the same function when a
string or a list is piped using the \c{\%>\%} function into the next expression.
The \c{NULL} arises from \c{R6::generator_funs} function when the \eval is
passed a \c{NULL} argument by the \c{DataMask_generator} package. The
\c{c("default", "default2012", "default2011" ...} pattern arises from
\c{copula::polyG} where it reflectively access the default expression for its
formal parameter and evaluates it. The \c{force(..1)} and \c{alist(`_spec`)}
patterns occur in \c{glue::glue_data} function which concatenates and
interpolates strings. The two patterns occur because the function captures
unevaluated unnamed arguments and maps the evaluation of \c{force(..1)} on them.
The \c{force} function forces promises~\cite{oopsla19a} and returns the result
of evaluation. The \c{inner} pattern arises from \c{glue::identity_transformer}
which enables the creation of custom transformation functions for affecting the
interpolation and concatenation of input by the \c{glue} package.

\subsection{Operation Mix}
In this section, we look at the operations executed by the expressions and
symbols on evaluation. R is a lazy language, so symbols can be bound to promises
that can evaluate arbitrary code. All operations in R are function calls, even
primitive operations such as addition and assignment are treated as calls to
functions implemented in the interpreter. In our corpus, we observe 2.9G calls
to R functions of which 125.5M happen inside \eval. Furthermore, unlike
Javascript, R code also calls into peformant native implementations of numerical
and statistical routines. In the corpus there are 463.8M calls to native code of
which only 31.2M happens inside \eval.
Qualitatively, we observe all kinds of operations being executed inside \eval
calls, from numerical computing for statistical modeling to reflection and
metaprogramming for implementing DSLs. Even testing frameworks such \c{genthat}
heavily rely upon \eval to execute test code snippets. The core R implementation
is rife with uses of \eval for mundane things like accessing default values of
formal parameters and in statistical packages for computation.

\subsubsection{Provenance}

In this section we turn our attention to the \AllExpressionInputEvalCallPerc
expressions passed as input to \eval. There are two ways to obtain
expressions in R.
\begin{itemize}
  \item \c{substitute} function can be used to synthesize ASTs from
  expressions by replacing symbols with their bindings in the specified
  environment.
  \item \c{expression} function can be used to create a vector of expression
    objects from text.
  \item \c{parse} function and its variants \c{str2expression} and \c{str2lang}
    can be used to turn strings or text from file into R expressions.
\end{itemize}

In our corpus, we observe 5.2\% cases where \eval directly evaluates the output of
\c{substitute}, 0.7\% cases where output of \c{parse} and its variants is read
directly and only 704 cases for \c{expression}. Majority of expression objects
consumed by \eval are generated by their transitive uses by other functions.

The pattern \c{eval(parse(...))} can be used for dynamic code loading. This
forms the core of \c{source} and \c{sys.source} functions in R that are commonly
used for loading code in R files in interactive settings. We investigated the
number of cases in which the output of \c{parse} and its variants is passed to
\eval, directly or transitively by tainting their output. This corresponds to \PercentParsedCallSites of the total \eval call sites and
\PercentParsedEvals of the eval calls. We observed that
very few of the eval call (\NbParseFilesRnd in total) consume the result of calling \c{parse} on
a file. Most of the eval calls consume the result of calling
\c{parse} on a string.  We also identified one
function in core R, \c{invokeRestartInteractively} that prompts the user for
input, parses it, and passes it to \eval.

%%Unlike Javascript, R packages also call native code for access to efficient
%%numerical and statistical routines. We intercept all calls to R functions and
%%native functions from the programs and \eval calls. In the whole corpus there
%%are \AllClosureCallCount calls to R functions and \AllNativeCallCount to native
%%code. Of those, \eval contribute to only \EvalClosureCallCount calls to R
%%functions and \EvalNativeCallCount to native code.
%%
%%\begin{table}[ht]\resizebox{\columnwidth}{!}{%
%%    \begin{tabular}{lll}\hline
%%      Source & Primitive Functions & R Functions & Native Functions \\ \hline
%%      Core & \CorePrimitiveFunctionCount & \CoreRFunctionCount & \CoreNativeFunctionCount \\
%%      CRAN & \PackagePrimitiveFunctionCount & \PackageRFunctionCount & \PackageNativeFunctionCount \\
%%      All  & \AllPrimitiveFunctionCount & \AllRFunctionCount & \AllNativeFunctionCount \\
%%    \end{tabular}} \label{table:operation-distribution}
%%  \caption{Distribution of operation calls to \eval calls across Core and Package}
%%\end{table}

@@

One of the main differences of the R \eval with the javascript one is how it
interacts with \emph{scope}, or environments in R. In Javascript, \c{eval} can
access the local scope and the global scope. In R, \c{eval} has an argument
\c{envir} to indicate in which environment its first argument must be evaluated.
By default, it is equal to \c{parent.frame()}, 
which is the parent environment of \c{eval} call. This default argument
happens in \DefaultEnvirExprPercent of the eval calls.

This argument can also be \c{NULL}, a \c{List} or a data frame. This happens in
3.5\% of the cases: \c{eval} copies the fields of the list or data frame and creates bindings for
them in a new environment. This pattern is used to evaluate formulas which can
directly refer to the fields of the data.
The \c{envir} argument can also be a number $n$. It means that the environment in
which the expressions is evaluated will be the result of \c{sys.call(n)} where
$n$ refers to the $n$-th stack frame.

@@

%%Below, we provide some code snippets to show how some of these environment
%%classes manifest.

%%The \c{match.arg} (see Listing~\ref{code:matchargenvir}) function has an
%%environment class 2. \c{sys.parent()} refers to the parent environment of
%%\c{match.arg} and so to the grand-parent environment of \c{eval} which we denote
%%as $2$.
%%
%%\begin{lstlisting}[label=code:matchargenvir, caption={\c{match.arg} using \eval with a custom environment.}]
%%formal.args <- formals(sys.function(sysP <- sys.parent()))
%%choices <- eval(formal.args[[as.character(substitute(arg))]], envir = sys.frame(sysP))
%%\end{lstlisting}

%%In package \emph{gamlss::get.K} function \c{get.K} evaluates its expression in its
%%parent environment, the one of the body of \c{get.K} returned from
%%\c{environment()}.
%%
%%\begin{lstlisting}
%%eval(parse(text = paste(\"fam$dld\", ii, sep = \"\")), envir = environment())
%%\end{lstlisting}
%%
%%
%%The \emph{glue} library is used to interpolate string litterals with variables, such as in \c{glue('My name is \{name\}.')} where \c{name} will be subtituted by its value. \emph{glue} makes it possible to define transformers that process those interpolated variables. They are built upon the \c{identity_transformer}, which uses \eval. The interpretation environment for \eval is often global here, as the string litterals are often used to print variables at the top level.
%%
%%\begin{lstlisting}
%%identity_transformer <- function(text, envir) 
%%{
%%    eval(parse(text = text, keep.source = FALSE), envir)
%%}
%%\end{lstlisting}
%%
%%
%% \end{lstlisting}
%% $$$$ <- to fix emacs fontification issues

The \c{imchange} function of package \c{imager} makes it possible to modify images using a dedicated formula syntax using \c{\~}.\footnote{Inspired by \c{map} in package \emph{purr}.}  Here, \eval is evaluated in \c{newenv}, which creates a new environment that inherits from \c{parent.frame()} by default (classified as 1+).

@@

The JavaScript paper identified a number of frequently occurring \eval patterns
which were detected using simple regular expressions on the \eval strings. In
the case of R, we match the patterns on the resolved AST.

@@

\begin{table}[ht]%\resizebox{\columnwidth}{!}{%
  \centering
\begin{tabular}{r|r|r|r|r}\hline
\multirow{2}{*}{Pattern} & \multicolumn{2}{c|}{Core} & \multicolumn{2}{c}{Packages} \\
                         & Calls & Ratio            & Calls & Ratio \\\hline
\input{tag/table-js-patterns-in-r.tex}
\end{tabular}
\caption{JavaScript evals patterns in R}
\label{table:js-pattens-in-r}
\end{table}

%\noindent
%\begin{compactitem}[$-$]
  % \item \emph{Library}: in JavaScript, the library pattern captures asynchronous
  % loading of external external function definitions. In R we have seen only
  % \PatternLibRnd such calls (\PatternLibRatio) mostly coming from source loading
  % functions (\ie \c{source} and \c{sys.source}). Most of the external function
  % definitions in R come from packages and R makes it very easy for users to
  % install packages from its centralized repository.

\noindent \emph{Library}, \emph{Typeof} and \emph{Try} are all barely used
mostly because the language provides dedicated features for it. The few calls in
Library pattern come mostly from file sourcing functions (\ie \c{source} and \c{sys.source}).
%
\noindent \emph{Read} pattern in JavaScript represented mostly an incorrect
access to an object's property or local variable both replaceable by JavaScript
core functions. In the case of R there are \PatternReadRnd read \eval calls
(\PatternReadRatio). They come from calls like \c{subset(airquality, select =
  Temp)} where the variable \c{Temp} will be evaluated in the \c{airquality}
data frame.
%
\noindent \emph{Assign} category comprises of assignments to variables within
an \eval call. We have seen \PatternAssignRnd such calls (\PatternAssignRatio).
\PatternAssignArrowRatio uses the simple assignment (\c{<-}). The majority of
cases comes from the \c{magrittr}'s \c{\%<>\%} pipe operator
(\PatternAssignArrowMagrittrRatio) which forwards an object to a function on rhs
and update lhs with the result. Far more calls used the super assignment
(\c{<<-}) operator, but almost all come from the \c{plyr} package which provides
a convenient API for common data manipulation problems.
%
\noindent \emph{Call} and \emph{Other} together represents the vast majority of
eval calls in R. The \emph{Other} pattern represents all the calls where eval
was called with a value instead of with an expression. The \emph{Call} pattern
contains all the calls that did not fall in the previous categories. Without a
context it is usually difficult to further categorize the intent of these calls.
Therefore we have randomly chosen \PatternManualPackages packages and manually
looked at \PatternManualCallsites call sites from which we have observed the following categories:

% this is read pattern
% \subsubsection{Evaluation using Data Fields} A common use of \eval is to
%   evaluate expressions that access data frame columns or list (heterogeneous
%   vectors) elements by name. Core R provides two functions, \c{with} and
%   \c{within} that implement this pattern. These functions construct a temporary
%   environment with data frame columns or list elements as the bindings.
%   Side-effects performed by the evaluated expression take place in this temporary
%   environment and are ignored by the \c{with} function. The \c{within} function
%   copies these modifications to a local copy of the input data and returns it.
%   This pattern is used quite often for statistical modeling. The modeling
%   expressions is constructed from variable names whose values are supplied by a
%   data frame or list.
\subsubsection{Reconding the current call} R provides a \c{match.call} function
which can return the current call. This is heavily used by the statistical
model-fitting functions that records the current call that is then reevaluated
in the different environment. Another pattern is to match the current call,
adjust some arguments and pass it to another function. For example the
\c{write.csv} function changes the \c{sep} argument in the matched call and reevaluates itself as \c{write.table}:
%
\begin{lstlisting}
    Call <- match.call(expand.dots = TRUE)
    Call$sep <- ","; Call[[1L]] <- as.name("write.table")
    eval.parent(Call)
\end{lstlisting}

@@

  \subsubsection{Miscellaneous} \eval is also used in R packages to get around
  various limitations. For example, in \datatable package, \eval is used to
  modify a binding in the \base package of R. Bindings in the \base package are
  locked for modification and require explicit unlocking using \unlockBinding
  function. However, directly using this function triggers a
  restricted-function-use warning during package build, preventing the package
  to be accepted by CRAN. The \datatable gets around this limitation by
  enclosing the call to \unlockBinding inside \eval. From the same reason, other
  packages use \eval to access private variables and functions from their
  dependencies.

  \subsubsection{Unnecessary use of eval} Similarly to JavaScript, there are also unnecessary uses of \eval.
  For example, the \c{PerfromanceAnalytics} package contains a function \c{chart.QQPlot} that uses \eval to resolve a string into function and anther to call it and assign its results into a variable:
  \begin{lstlisting}
function (R, d="norm", dp, ...) {
q.f <- eval(parse(text=paste("q",d,sep="")))
z <- NULL
eval(parse(text=paste("z<-q.f(",dp,",...)")))
}
  \end{lstlisting}
  In both cases, there is no need for \eval:
  \begin{lstlisting}
function (R, d="norm", dp, ...) {
q.f <- get(paste0("q",d))
z <- q.f(dp, ...)
}
  \end{lstlisting}
  or even to a oneliner \c{do.call(paste0("q",d), as.list(dp, ...))}.

@@

%\begin{figure}[!h]
%	\centering
%    \includegraphics[width=\columnwidth]{polymorphism}
%    \caption{Number of different types  of \c{expr} per call sites.} \label{fig:polymorphism}
%\end{figure}
%
%\subsection{base}
%
%%There are 30 functions in \emph{base} that use \eval. Only five of them use \c{parse}, for legitimate reasons, such \c{source} and \c{sys.source} which load a file and then evaluate it, or \c{invokeRestartInteractively} which prompts the user for arguments to restart execution after an error.
%%
%%We endeavored  to write these functions without \eval. \c{match.arg} appears in \MatchArgPercent of the \eval calls. It is used to get the arguments, possibly abbreviated, of a function. With one argument, as \c{match.arg(arg)}, it uses \c{eval} to infer them from a parent function's formal arguments. We can rewrite it using its version with two arguments, \c{match.arg(arg, choices)} where \c{choices} are the possible formal arguments of the function, as shown in Listing~\ref{code:matcharg}.
%%
%%\begin{lstlisting}[caption={Rewriting \c{match.arg} without \eval}, label=code:matcharg]
%%f <- function(type=c("examples", "tests", "vignettes")) {
%%actual_type = match.arg(type)
%%...
%%}
%%
%%f <- function(type=c("examples", "tests", "vignettes")) {
%%actual_type = match.arg(type, choices=c("examples", "tests", "vignettes"))
%%...
%%}
%%\end{lstlisting}

>>>>

This section presents an overview of the usage of \eval in our corpus. Richards
et al. reported 82\% of JavaScript-enabled web pages using \eval~\cite{ecoop11}.
The case for R is more striking with 100\% of the programs call the \eval
function.\footnote{We combine calls to {\sf eval}, {\sf eval.parent}, {\sf
    evalq} and {\sf local}.} This high number is explained by the fact that the
R implementation, its core packages, use \eval extensively. We differentiate
three types of eval based on their origin:
%
\begin{compactitem}[$-$]
  \item a {\bf core eval} is one that occurs in the code of a core package,
  \item a {\bf dependent eval} occurs in a package imported by the target
  package, and
  \item an {\bf own eval} is one that occurs in the source code of the target
  package.
\end{compactitem}

@@

%% SORRY  I HAD TO REFORMAT
\begin{table}[!h] \centering
 \begin{tabular}{r|c|c|c} \hline
Source & Core    & Dependent                      & Own \\\hline
CRAN   & 5.6M &  867K    & 259K \\
       & (1 / 337 / 46K)& (0 / 52 / 58K ) & (0 / 15 / 15K) \\\hline
Kaggle &  1.2M &  470K   & 0\\
 &  (0 / 721 / 18K) & (0 / 290 / 20K) & (0 / 0 / 0)\\\hline
\end{tabular}\caption{Eval calls in CRAN and Kaggle}\label{A}\end{table}

Table~\ref{table:site-package-summary} shows the number of call sites in
CRAN packages. \OneCallSitesEnrich{} packages have a single call site, and one
package, \emph{metafor}, has \MaxCallSitesEnrich{} call sites of
\eval. \NoCallSitesEnrich{} packages have no calls to \eval. The core packages
have \NbCoreEvalCallSites{} call sites of \eval and Kaggle has none. These
numbers are under-approximations as \eval may be called reflectively,
through aliases, or passed to a higher-order function.

@@

In this section we turn our attention to the expression passed for
evaluation to \eval. The \eval function accepts all types of R objects as
inputs, but in the majority of cases (\AllExpressionInputEvalCallPerc), it
is called with expressions. The remaining \AllValueInputEvalCallPerc are
values, \ie, they evaluate to themselves.
%
Passing a value into an eval is effectively an no-op so we look closely at the
expression types the call sites pass to eval across all the runs.
\AllValueInputEvalSitePerc of all call sites always pass values,
\AllExpressionInputEvalSitePerc pass expressions and a very large proportion,
\AllPolymorphicInputEvalSitePerc, pass both expressions and values as inputs
across all program runs. We inspected the eval callsites that accept only values
or both values and expression as inputs. The top three of these call sites
contribute to over ~80\% of all calls to evals. The first one is inside
\c{match.arg} from core. Given a vector of names as strings, this function finds
the default values for parameters with those names in the caller function. Since
the default value for a function parameter can be a complex expression or a
benign value such as \c{NULL}, the \c{match.arg} function evaluates it through a
call to \eval. This makes the \eval inside \c{match.arg} ``polymorhpic''. The
second such \eval call originates from \ggplot, a popular R package for making
plots. \ggplot has it's own object system, \ggproto, which uses \eval inside the
\ggproto function to determine the super-class from which it inherits (the
default value is \c{NULL} but in general it could be an arbitrarily expression).
The third \eval call is in core R \c{str} function used to print a summary of R
objects. It internally maps \eval function on the bindings of an environment
which results in inputs of multiples types being passed to \eval.

@@

We looked at the top ten expressions passed to core and package \eval
calls. The most frequent ten expressions to eval calls from core R
contribute to 85\% of all \eval calls.

\begin{table}[!h] \centering
\begin{tabular}{@{}l|rr@{}} \hline
Expression & Eval Call &  \% \\\hline
\c{c("auto", "shell", "radix")} & 198,7105 & 29\%\\
\c{c("auto", "shell", "quick", "radix")} & 1,593,169  & 23\%\\
\c{\{info <- loadingNamespaceInfo(...} & 1,008,632 &       14\%\\
\c{c("onLoad", "attach", "detach", "onUnload")}   & 470,566 &      6.9\%\\
\c{c("append", "prepend", "replace")} &              261,587&       3.9\% \\
\c{c("left", "right", "centre", "none")} & 162,086     & 2.4\%\\
\c{c("no", "ifany", "always")}   &                71,580 &       1.1\%\\
\c{c("pearson", "kendall", "spearman")}  & 72,962 &      1.1\%\\
\c{NULL}& 75330  &      1.1\% \\
\c{Symbol}&                 6,6279&       1\%\\\hline
\end{tabular}
\caption{Top ten eval calls in Core}\label{B}
\end{table}

@@

The most frequent ten expressions to eval calls from CRAN packages
contribute to 77.1\% of all \eval calls.

\begin{table}[!h]  \centering
\begin{tabular}{@{}l@{~}|@{~}r@{~}r@{}} \hline
Expression & Eval Call &  \% \\\hline
\c{Environment} &                                  989302   & 61\%\\
\c{column[rows] <<- what} &                        55677    & 3.5\%\\
\c{function(value) freduce(value, `_function_list`)} & 37251& 2.3\%\\
\c{NULL} &                         32005    & 2\%\\
\c{List} &                         22293    & 1.4\%\\
\c{c("default", "default2012", "default2011" ...}& 20610    & 1.3\%\\
\c{force(..1)}            &                        20461    & 1.3\%\\
\c{alist(`_spec`)}       &                         18532    & 1.2\%\\
\c{inner}               &                          18530    & 1.2\%\\
\c{String Vector}      &                           17487     & 1.1\%\\
\end{tabular}\caption{Top ten eval calls in CRAN} \label{C}
\end{table}

The expression \c{Environment} occurs because of the four callsites
explained above, \c{ggplot2::ggproto}, \c{R6::generator_funs},
\c{future::backtrace} and \c{RModel::str.RMmodel}. The next expression,
\c{column[rows] <<- what}, is used inside the \c{plyr::rbind.fill} function
to merge data frames by assigning concatenated vectors to rows. The \c{<<-}
operator is interesting in that it skips the current scope and assigns in a
parent scope in which the variable is already present. In our corpus, all
these \eval calls contribute to a single side-effect. The expression
\c{function(value) freduce(value, `_function_list`)} arises from the
\c{magrittr::\%>\%} function which is a pipe operator that pipes the output
of previous command to the next one. The expression is evaluated in a custom
environment to create a function binding for evaluating the components of
the pipe.\c{String Vector} and \c{List} also arise from the same function
when a string or a list is piped using the \c{\%>\%} function into the next
expression.  The \c{NULL} arises from \c{R6::generator_funs} function when
the \eval is passed a \c{NULL} argument by the \c{DataMask_generator}
package. The \c{c("default", "default2012", "default2011" ...} pattern
arises from \c{copula::polyG} where it reflectively access the default
expression for its formal parameter and evaluates it. The \c{force(..1)} and
\c{alist(`_spec`)} patterns occur in \c{glue::glue_data} function which
concatenates and interpolates strings. The two patterns occur because the
function captures unevaluated unnamed arguments and maps the evaluation of
\c{force(..1)} on them.  The \c{force} function forces promises and returns
the result of evaluation. The \c{inner} pattern arises from
\c{glue::identity_transformer} which enables the creation of custom
transformation functions for affecting the interpolation and concatenation
of input by the \c{glue} package.

\subsection{Operation Mix}

The JavaScript paper reported on the operation mix within \evals. This means
the proportion of read, write, and call bytecodes performed by the
JavaScript interpreter. They found that some web pages were creating objects
more frequently and writing into their fields.  In R, nearly every operation
performed in the interpreter is a function call. We observe 2.9G calls and,
of those 125M happen inside \eval (there are 463M native calls, of which
only 31M are in \eval).

{\bf there is no good reason for not counting reads and writes, we can...}

%% ...The following is fluff ...

%% Qualitatively, we observe all kinds of operations being executed inside
%% \eval calls, from numerical computing for statistical modeling to reflection
%% and metaprogramming for implementing DSLs. Even testing frameworks such
%% \c{genthat} heavily rely upon \eval to execute test code snippets. The core
%% R implementation is rife with uses of \eval for mundane things like
%% accessing default values of formal parameters and in statistical packages
%% for computation.

\subsection{Provenance}

In JavaScript, the provenance of the string passed into \eval is a matter of
concern as \eval can be used to breach the security of websites. The main
danger comes from string that were obtained from external sources. If they
are allowed to flow into an \eval, then an attacker could hijack the
program.

{\bf what have they found...}

There are various ways to obtain expressions:
\begin{itemize}
  \item \c{substitute} synthesizes ASTs from expressions by replacing
    symbols with their bindings in the specified environment.
  \item \c{expression}  creates a vector of expression
    objects from text.
  \item \c{parse}, \c{str2expression} and \c{str2lang} turn strings into
    expressions.
\end{itemize}

We observe 5.2\% cases where \eval directly evaluates the output of
\c{substitute}, 0.7\% cases where output of \c{parse} is read directly and
only 704 cases for \c{expression}. Most expressions consumed by \eval are
generated by other functions.

\c{eval(parse(...))} can be used for dynamic code loading. This forms the
core of \c{source} and \c{sys.source} functions in R that are commonly used
for loading code in R files in interactive settings. We investigated the
number of cases in which the output of \c{parse} and its variants is passed
to \eval, directly or transitively by tainting their output. This
corresponds to \PercentParsedCallSites of the total \eval call sites and
\PercentParsedEvals of the eval calls. We observed that very few of the eval
call (\NbParseFilesRnd in total) consume the result of calling \c{parse} on
a file. Most of the eval calls consume the result of calling \c{parse} on a
string.  We also identified one function in core R,
\c{invokeRestartInteractively} that prompts the user for input, parses it,
and passes it to \eval.

@@

One of the main differences of the R \eval with the javascript one is how it
interacts with \emph{scope}, or environments in R. In Javascript, \eval can
access the local scope and the global scope. In R, \eval has an argument
\c{envir} to indicate in which environment its first argument must be evaluated.
By default, it is equal to \c{parent.frame()}, 
which is the parent environment of \eval call. This default argument
happens in \DefaultEnvirExprPercent of the eval calls.

This argument can also be \c{NULL}, a \c{List} or a data frame. This happens
in 3.5\% of the cases: \eval copies the fields of the list or data frame
and creates bindings for them in a new environment. This pattern is used to
evaluate formulas which can directly refer to the fields of the data.  The
\c{envir} argument can also be a number $n$. It means that the environment
in which the expressions is evaluated will be the result of \c{sys.call(n)}
where $n$ refers to the $n$-th stack frame.

@@

The \c{imchange} function of package \c{imager} makes it possible to modify
images using a dedicated formula syntax using \c{\~}.\footnote{Inspired by
 {map} in package \emph{purr}.}  Here, \eval is evaluated in \c{newenv},
which creates a new environment that inherits from \c{parent.frame()} by
default (classified as 1+).

@@

Richards et al.~\cite{ecoop11} identified a number of frequently occurring \eval
patterns which were detected using simple regular expressions on the \eval
strings. In the case of R we match the patterns on the recorded AST.

@@

\begin{table}[ht]%\resizebox{\columnwidth}{!}{%
  \centering
\begin{tabular}{r|r|r|r|r}\hline
\multirow{2}{*}{Pattern} & \multicolumn{2}{c|}{Core} & \multicolumn{2}{c}{Packages} \\
                         & Calls & Ratio            & Calls & Ratio \\\hline
\input{tag/table-js-patterns-in-r.tex}
\end{tabular}
\caption{JavaScript \eval Patterns in R}
\label{table:js-pattens-in-r}
\end{table}

\noindent \emph{Library}, \emph{Typeof} and \emph{Try} are all barely used
mostly because the language provides dedicated features for it. The few calls in
Library pattern come mostly from file loading functions such as \c{source} and
\c{sys.source}.
%
\noindent \emph{Read} pattern in JavaScript represented mostly an incorrect

access to an object's property or local variable both replaceable by
JavaScript core functions. In the case of R there are \PatternReadRnd read
\eval calls (\PatternReadRatio). They come from calls like
\c{subset(airquality, select = Temp)} where the variable \c{Temp} will be
evaluated in the \c{airquality} data frame.

%
\noindent \emph{Assign} category comprises of assignments to variables
within an \eval call. We have seen \PatternAssignRnd such calls
(\PatternAssignRatio).  \PatternAssignArrowRatio uses the simple assignment
(\c{<-}). The majority of cases comes from the \c{magrittr}'s \c{\%<>\%}
pipe operator (\PatternAssignArrowMagrittrRatio) which forwards an object to
a function on rhs and update lhs with the result. Far more calls used the
super assignment (\c{<<-}) operator, but almost all come from the \c{plyr}
package which provides a convenient API for common data manipulation
problems.
%
\noindent \emph{Call} and \emph{Other} together represents the vast majority
of eval calls in R. The \emph{Other} pattern represents all the calls where
eval was called with a value instead of with an expression. The \emph{Call}
pattern contains all the calls that did not fall in the previous
categories. Without a context it is usually difficult to further categorize
the intent of these calls.  Therefore we have randomly chosen
\PatternManualPackages packages and manually looked at
\PatternManualCallsites call sites from which we have observed the following
categories:

\subsubsection{Recording the current call} R provides a \c{match.call} function
which can return the current call. This is heavily used by the statistical
model-fitting functions that records the current call that is then
reevaluated in the different environment. Another pattern is to match the
current call, adjust some arguments and pass it to another function. For
example the \c{write.csv} function changes the \c{sep} argument in the
matched call and reevaluates itself as \c{write.table}:
%
\begin{lstlisting}
    Call <- match.call(expand.dots = TRUE)
    Call$sep <- ","; Call[[1L]] <- as.name("write.table")
    eval.parent(Call)
\end{lstlisting}

@@

  \subsubsection{Miscellaneous} \eval is also used in R packages to get around
  various limitations. For example, in \datatable package, \eval is used to
  modify a binding in the \base package of R. Bindings in the \base package are
  locked for modification and require explicit unlocking using 
  function \unlockBinding. However, directly using this function triggers a
  restricted-function-use warning during package build, preventing the package
  to be accepted by CRAN. The \datatable gets around this limitation by
  enclosing the call to \unlockBinding inside \eval. From the same reason, other
  packages use \eval to access private variables and functions from their
  dependencies.

  \subsubsection{Unnecessary use of eval}
  Similarly to JavaScript, there are also unnecessary uses of \eval.
  For example, the \c{PerformanceAnalytics} package contains a function
  \c{chart.QQPlot} that uses \eval to resolve a string into function and
  another to call it and assign its results into a variable:
  \begin{lstlisting}
function (R, d="norm", dp, ...) {
q.f <- eval(parse(text=paste("q",d,sep="")))
z <- NULL
eval(parse(text=paste("z<-q.f(",dp,",...)")))
}
  \end{lstlisting}
  In both cases, there is no need for \eval:
  \begin{lstlisting}
function (R, d="norm", dp, ...) {
q.f <- get(paste0("q",d))
z <- q.f(dp, ...)
}
  \end{lstlisting}
  or even to a oneliner \c{do.call(paste0("q",d), as.list(dp, ...))}.

@@

\section{Threats to Validity} The primary threat to validity is the issue of
code coverage, a cause of concern for any dynamic approach. We mitigate this
by including focusing on R packages that have high code coverage. We turn
off the bytecode compiler for this study. The bytecode compiler can also
call \eval. We do not get source locations for \UndefinedEvalsRnd eval
calls. In these cases \eval is either passed as an argument to a
higher-order functions or is defined in a function returned by a
higher-order function and the R parser does not retain location information
for \eval. However, this is a meager \PercentUndefinedEval of all eval calls
and is unlikely to affect our analysis. We ignore calls to the native \eval
function exposed by R. We also ignore the \c{rlang::tidy_eval} function
which uses native \eval internally because \c{rlang} is used to implement a
DSL for data analysis in R. It introduces a new first-class promise object
called \c{quosure} for which it implements special evaluation support in
\c{tidy_eval}.
